---
title: "Open Research Data Program of the ETH Board"
subtitle: "proposal for an open metadata dataset"
author: 
  - name: Lars Schöbitz
    affiliations:
        - name: Global Health Engineering - ETH Zurich
          url: https://ghe.ethz.ch/
date: 2025-05-12
date-format: long
format: 
  revealjs:
    theme: slides.scss
    transition: fade
    slide-number: true
    width: 960
    height: 540
editor: source
editor_options: 
  chunk_output_type: console
execute:
  freeze: auto
---

```{r}
library(tidyverse)
library(ggthemes)
```

```{r}
data <- read_csv("https://raw.githubusercontent.com/Global-Health-Engineering/ethord/refs/heads/main/data/data.csv") |> 
  janitor::clean_names()
```

```{r}
df_category <- data |> 
  count(category)

n_projects <- sum(df_category$n)

n_contribute <-  filter(df_category, category == "Contribute")$n
n_explore <-  filter(df_category, category == "Explore")$n
n_establish <-  filter(df_category, category == "Establish")$n
```

## Measure 1: Calls for Field-Specific Actions

- 11 calls for proposals
- `r n_projects` funded projects (`r n_contribute` Contribute, `r n_explore` Explore, `r n_establish` Establish)
- X.X million CHF in funding

::: footer
Data from: 
:::

## Program evaluation {.scrollable .smaller}

::: incremental

- Who are the researchers that received funding? {{< fa thumbs-up >}} 
- Which institutions are they employed at? {{< fa thumbs-up >}}
- Which disciplines do they represent? {{< fa thumbs-up >}} **kind of**

:::

##

```{r}
#| out-width: "100%"
#| fig-width: 10
#| fig-height: 5

data |> 
  count(institutions, category) |> 
  ggplot(aes(x = reorder(institutions, n), 
             y = n,
             fill = category)) +
  geom_col(position = "dodge") +
  geom_label(aes(label = n),
             position = position_dodge(width = 0.9),
             show.legend = FALSE,
             color = "white",
             fontface = "bold",
             size = 3) +
  coord_flip() +
  labs(
    title = "Open Research Data Program of the ETH Board",
    subtitle = "Number of funded projects per institution and project category",
    y = "Number of projects",
    x = NULL,
    fill ="Project category:"
  ) +
  scale_fill_colorblind() +
  statR::theme_stat(base_size = 10) +
  theme(panel.grid = element_blank(),
        axis.text.y = element_text(size = 8)) 

```

::: footer

Data from: <https://raw.githubusercontent.com/Global-Health-Engineering/ethord/refs/heads/main/data/data.csv>

:::

## 



## Program evaluation {.scrollable .smaller}

::: incremental

- Who are the researchers that received funding? {{< fa thumbs-up >}} 
- Which institutions are they employed at? {{< fa thumbs-up >}}
- Which disciplines/fields do they represent?  {{< fa thumbs-up >}} **kind of**
- What is their scientific role (e.g. Professor, Post Doc, PhD, etc.)? {{< fa thumbs-down >}}
- How many publications are derived from these projects? {{< fa thumbs-down >}}
- How many ORD datasets have been published? {{< fa thumbs-down >}}
- How where budgets distributed among their cost categories? {{< fa thumbs-down >}}
- How many proposals were submitted to each call? And how many received funding? {{< fa thumbs-down >}}
- etc.

:::

## Problem statement

A metadata dataset that brings together all the information from the ETH ORD programme is currently not available. 

## Proposed solution

Project relevant internal data sources (e.g. submitted proposals, final scientific reports, MS Excel files, Email exchange, online research) will be used to prepare a set of metadata datasets. 

## Make it Open and FAIR

In the spirit of ORD, the metadata datasets will be shared as open research data, following FAIR principles for data sharing.

- permissive license (CC0)
- assigned DOIs
- rich documentation
- various formats (e.g. CSV, XLSX, R data package)
- curated (stand-alone website)
- etc.

::: speaker-notes
- Findable
- unique identifiers (e.g. DOIs)
- metadata
- persistent URLs
- Accessible
- open access
- no login required
- no restrictions on use
- Interoperable
- machine-readable
- standard formats (e.g. CSV, JSON, XML)
- standard vocabularies (e.g. schema.org)
- Reusable
- clear usage rights
- clear provenance
- clear licensing
- clear attribution
:::

## SNFS Grants database

Not FAIR, but open: 

::: learn-more
[data.snf.ch/stories/snsf-open-data-who-gets-funding-en.html](https://data.snf.ch/stories/snsf-open-data-who-gets-funding-en.html)

```{=html}
<iframe src="https://data.snf.ch/stories/snsf-open-data-who-gets-funding-en.html" width="100%" height="500" style="border:2px solid #123233;" title="SNSF Grants database"></iframe>
```
:::

# Datasets

At minimum, the following datasets are suggested, which can be combined by unique identifier (proposal_id):

## Dataset: `proposals` {.scrollable .smaller}

- call_id
- call_category
- proposal_id
- proposal_url (if proposal is public somehere)
- acronym
- title
- abstract
- funding_requested
- funding_received
- panel_score_excellence_reviewer1
- panel_score_implementation_reviewer1
- panel_score_impact_reviewer1
- panel_score_excellence_reviewer2
- panel_score_implementation_reviewer2
- panel_score_impact_reviewer2
- acceptance (yes/no)
- main_applicant_name (if too for rejected proposals, remove from ORD data, but keep internal resource)
- main_applicant_institution_name
- main_applicant_department_name
- main_applicant_lab_name
- co_applicant_name (if too for rejected proposals, remove from ORD data, but keep internal resource)
- co_applicant_institution_name
- co_applicant_department_name
- co_applicant_lab_name

## Dataset: `proposals` - Sources

- Project proposals (accepted & rejected)
- Internal MS Excel files for scores
- Email exchange
- Online research

**Note:** If access to rejected proposals is not possible, a fourth dataset `calls` is suggested.

## Dataset: `projects` {.scrollable .smaller}

- proposal_id
- project_start
- project_end
- no_cost_extension
- project_website
- budget_cost_category_senior_staff
- budget_cost_category_postdocs
- budget_cost_category_students
- budget_cost_category_other
- budget_cost_category_publications
- budget_cost_category_conferences_workshops
- budget_cost_category_other
- budget_sub_contracting_costs
- etc.

## Dataset: `projects` - Sources

- Project proposals
- Scientific reports

## Dataset: `outputs` {.scrollable .smaller}

- proposal_id
- category (e.g. website, publication, dataset, software, etc.)
- repository_catalogue
- etc.

## Dataset: `outputs` - Sources

- Scientific reports, list of outputs
- Email exchange

## Dataset: `calls` {.scrollable .smaller}

A summary table:

- call_id
- call_category
- outcome_total (e.g. accepted, rejected)
- value

## Thanks! 

Slides created via revealjs and Quarto: <https://quarto.org/docs/presentations/revealjs/> 

Data ([Nicoló Massari](https://github.com/massarin)) available at: <https://raw.githubusercontent.com/Global-Health-Engineering/ethord/refs/heads/main/data/data.csv>

Access slides as [PDF on GitHub](https://github.com/global-health-engineering/website/raw/main/slides/eth-ord-data.pdf)

